---
title: "BMP treatments"
output: html_document
fig_caption : yes
theme : cerulean
toc : yes
toc_depth : 3
toc_float : yes
---

# 1. Download packages

```{r}

options(timeout = 1000)
library("patchwork")
library("forcats")
library ("multcompView")
library("ggplot2") 
library("Hmisc")
library("tidyverse")
library("dplyr")
library("RColorBrewer")
library("metacoder")
library("patchwork")
library("gridExtra")
library("microeco")
library ("MASS")
install.packages("randomForest", repos = "http://cran.us.r-project.org")
library ("randomForest") 
library ("ggdendro")
library ("ggrepel")
library ("agricolae")
library ("igraph")
library ("picante")
library ("pheatmap")
library ("rgexf")
library ("ggalluvial")
library ("ggh4x")
library ("rcompanion")
library ("FSA")
library ("gridExtra")
library ("aplot")
library ("NST")
library ("GGally")
library ("ggraph")
library ("networkD3")
library ("poweRlaw")
library ("ggtern")
library ("SRS")
library
library("metan")
library("htmltools")



```

# 2. Import the data
## 2.1. Anova data

```{r}
file1 <- 'BMP-ANOVA.csv'
BMP1 <- read.csv(file1, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
str(BMP1)
BMP_sludge <- BMP1[1:15,]
BMP_MCC <- BMP1[16:30,]

```

## 2.2. Data for microbial analysis
```{r}
file2 <- 'RelativeAbundanceBMP.csv'
Data1 <- read.csv(file2, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
file3 <- 'Taxonomy.csv'
Data2 <- read.csv(file3, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
file4 <- 'CaracterizationBMP.csv'
Data3 <- read.csv(file4, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)

file5 <- 'Abundancetaxonomy.csv'
Data4 <- read.csv(file5, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
```
## 2.3. Data for correlation analysis
```{r}
file6 <- 'CorrelationBMPMCC.csv'
MCC <- read.csv(file6, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
file7 <- 'CorrelationBMPSS.csv'
SS <- read.csv(file7, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
```
# 3. ANOVA analysis
## 3.1. Anova BMP
```{r}
ANOVAsludge <- aov(BMP ~ Temperature * factor(Inoculum), data=BMP_sludge)
  summary(ANOVAsludge)
  ANOVAsludge$coefficients
  
ANOVAMCC <- aov(BMP ~ Temperature * factor(Inoculum), data=BMP_MCC)
  summary(ANOVAMCC)
```
## 3.2. Anova t80
```{r}
ANOVAsludge <- aov(t80 ~ Temperature * factor(Inoculum), data=BMP_sludge)
summary(ANOVAsludge)

ANOVAMCC <- aov(t80 ~ Temperature * factor(Inoculum), data=BMP_MCC)
summary(ANOVAMCC)

```
## 3.3. Tukey t80
```{r}
TukeySludge <- TukeyHSD(ANOVAsludge)
print (TukeySludge)

TukeyMCC <- TukeyHSD(ANOVAMCC)
print (TukeyMCC)

TukeySludge.cld <- multcompLetters4(ANOVAsludge,TukeySludge)
print(TukeySludge.cld)

TukeyMCC.cld <- multcompLetters4(ANOVAMCC,TukeyMCC)
print(TukeyMCC.cld)
```
# 3. Microbial analysis
## 3.1. Bacterial abundance at Phylum level
```{r}
dat2 <- Data1 %>%
  pivot_longer(-X.OTU.ID, names_to = "Sample_ID", values_to = "Value")
dat2

dat2 <- dat2 %>%
  left_join(Data2, by = "X.OTU.ID")
dat2

dat2 <- dat2 %>%
  left_join(Data3, by = "Sample_ID")
dat2
dat2[dat2==0] <- NA
datcom2 <-dat2[complete.cases(dat2),]

nb.cols2 <- 30
mycolors2 <- colorRampPalette(brewer.pal(30, "Paired"))(nb.cols2)
datcom2 %>%
  ggplot(aes(x = Sample_ID, y = Value)) + 
  geom_bar(aes(fill= Phylum), stat = "identity", position = "fill")+
  scale_fill_manual(values = mycolors2) +
  xlab("Sample") + ylab("Relative abundance")+
  facet_grid(~ Origin, scales = "free_x", space = "free_x")
```
## 3.2. Heat trees at Genus level
## 3.2.1. Data preparation

```{r}
OTU_dat <- left_join(Data4, Data2,
                                 by = c("X.OTU.ID" = "X.OTU.ID"))

OTU_datI1MES <- OTU_dat[,-c(3:6)]
OTU_datI1MES[OTU_datI1MES==0] <- NA
OTU_datI1MEScom <-OTU_datI1MES[complete.cases(OTU_datI1MES),]

OTU_datI2MES <- OTU_dat[,-c(2,4:6)]
OTU_datI2MES[OTU_datI2MES==0] <- NA
OTU_datI2MEScom <-OTU_datI2MES[complete.cases(OTU_datI2MES),]

OTU_datI2THER <- OTU_dat[,-c(2,3,5,6)]
OTU_datI2THER[OTU_datI2THER==0] <- NA
OTU_datI2THERcom <-OTU_datI2THER[complete.cases(OTU_datI2THER),]

OTU_datI3MESO <- OTU_dat[,-c(2:4,6)]
OTU_datI3MESO[OTU_datI3MESO==0] <- NA
OTU_datI3MESOcom <-OTU_datI3MESO[complete.cases(OTU_datI3MESO),]

OTU_datI3THER <- OTU_dat[,-c(2:5)]
OTU_datI3THER[OTU_datI3THER==0] <- NA
OTU_datI3THERcom <-OTU_datI3THER[complete.cases(OTU_datI3THER),]
```
## 3.2.2. Data parsing

```{r}
datataxonomyI1MES <- parse_tax_data(OTU_datI1MEScom,
                                 class_cols = "taxonomy",
                                 class_sep = ";",
                                 class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
                                 class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))

datataxonomyI2MES <- parse_tax_data(OTU_datI2MEScom,
                                 class_cols = "taxonomy",
                                 class_sep = ";",
                                 class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
                                 class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))

datataxonomyI2THER <- parse_tax_data(OTU_datI2THERcom,
                                 class_cols = "taxonomy",
                                 class_sep = ";",
                                 class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
                                 class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))

datataxonomyI3MESO <- parse_tax_data(OTU_datI3MESOcom,
                                 class_cols = "taxonomy",
                                 class_sep = ";",
                                 class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
                                 class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))

datataxonomyI3THER <- parse_tax_data(OTU_datI3THERcom,
                                 class_cols = "taxonomy",
                                 class_sep = ";",
                                 class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
                                 class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))


names(datataxonomyI1MES$data) <- "OTU_counts"
names(datataxonomyI2MES$data) <- "OTU_counts"
names(datataxonomyI2THER$data) <- "OTU_counts"
names(datataxonomyI3MESO$data) <- "OTU_counts"
names(datataxonomyI3THER$data) <- "OTU_counts"
```

## 3.2.2. Heat trees visualization
```{r}
datataxonomyI1MES %>% 
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "g", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = gsub(pattern = "\\[|\\]", replacement = "", taxon_names),
            node_size = n_obs,
            node_color = n_obs,
            node_color_axis_label = "OTU count",
            title = "I1_meso",
            layout = "davidson-harel", initial_layout = "reingold-tilford")

datataxonomyI2MES %>% 
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "g", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = gsub(pattern = "\\[|\\]", replacement = "", taxon_names),
            node_size = n_obs,
            node_color = n_obs,
            node_color_axis_label = "OTU count",
            title = "I2_meso",
            layout = "davidson-harel", initial_layout = "reingold-tilford")

datataxonomyI2THER %>% 
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "g", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = gsub(pattern = "\\[|\\]", replacement = "", taxon_names),
            node_size = n_obs,
            node_color = n_obs,
            node_color_axis_label = "OTU count",
            title = "I2_thermo",
            layout = "davidson-harel", initial_layout = "reingold-tilford")

datataxonomyI3MESO %>% 
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "g", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = gsub(pattern = "\\[|\\]", replacement = "", taxon_names),
            node_size = n_obs,
            node_color = n_obs,
            node_color_axis_label = "OTU count",
            title = "I3_meso",
            layout = "davidson-harel", initial_layout = "reingold-tilford")

datataxonomyI3THER %>% 
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "g", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = gsub(pattern = "\\[|\\]", replacement = "", taxon_names),
            node_size = n_obs,
            node_color = n_obs,
            node_color_axis_label = "OTU count",
            title = "I3_thermo",
            layout = "davidson-harel", initial_layout = "reingold-tilford")
```
# 3.3. Correlation analysis
## 3.3.1. Alpha diversity indices

```{r}
Data5 <- Data4[,-c(1,7)]
rownames(Data5) <- Data4[,1]
Data6 <- Data3
rownames(Data6) <- Data3[,1]
Data7 <- Data2[,-1]
rownames(Data7) <- Data2[,1]

dataset <- microtable$new(sample_table = Data6, otu_table = Data5, tax_table = Data7)

dataset$rarefy_samples(sample.size = 1500)

dataset$cal_abund()

dataset$cal_alphadiv(PD = FALSE)
dataset$alpha_diversity

```

## 3.3.1. Correlation between relative abundance and inoula caracteristics
```{r}
t2 <- trans_diff$new(dataset = dataset, method  = "metastat",group = "Sample_ID", alpha = 0.1,
                     
                                       p_adjust_method ="none", taxa_level = "Genus")
t1 <- trans_env$new(dataset = dataset, add_data = Data6[,-c(13:16)])
t1$cal_cor(use_data = "other", p_adjust_method = "fdr",
           other_taxa = t2$res_diff$Taxa[1:60], cor_method = "pearson")
t1$plot_cor()

```

## 3.3.1. Correlation between relative abundance and inoula caracteristics
```{r}
MCC1 <- MCC [,-c(1,2)] 
SS1 <- SS[,-c(1,2)]
corrMCC <- corr_coef(MCC1)
corrSS <- corr_coef(SS1)
plot (corrMCC)
plot (corrSS)
```